### 2010-11-09 journal/paper.tex
Start adding agda to the paper

\documentclass[a4paper]{article}

%% ODER: format ==         = "\mathrel{==}"
%% ODER: format /=         = "\neq "
%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stmaryrd}
\DeclareFontFamily{OT1}{cmtex}{}
\DeclareFontShape{OT1}{cmtex}{m}{n}
  {<5><6><7><8>cmtex8
   <9>cmtex9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
\DeclareFontShape{OT1}{cmtex}{m}{it}
  {<-> ssub * cmtt/m/it}{}
\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
  {<5><6><7><8>cmtt8
   <9>cmbtt9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
\DeclareFontShape{OT1}{cmtex}{bx}{n}
  {<-> ssub * cmtt/bx/n}{}
\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU

\newcommand{\Sp}{\hskip.33334em\relax}

\newcommand{\Conid}[1]{\mathit{#1}}
\newcommand{\Varid}[1]{\mathit{#1}}
\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
\newcommand{\plus}{\mathbin{+\!\!\!+}}
\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
\newcommand{\sequ}{\mathbin{>\!\!\!>}}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\usepackage{polytable}

%mathindent has to be defined
\@ifundefined{mathindent}%
  {\newdimen\mathindent\mathindent\leftmargini}%
  {}%

\def\resethooks{%
  \global\let\SaveRestoreHook\empty
  \global\let\ColumnHook\empty}
\newcommand*{\savecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
\newcommand*{\restorecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
\newcommand*{\aligncolumn}[2]%
  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}

\resethooks

\newcommand{\onelinecommentchars}{\quad-{}- }
\newcommand{\commentbeginchars}{\enskip\{-}
\newcommand{\commentendchars}{-\}\enskip}

\newcommand{\visiblecomments}{%
  \let\onelinecomment=\onelinecommentchars
  \let\commentbegin=\commentbeginchars
  \let\commentend=\commentendchars}

\newcommand{\invisiblecomments}{%
  \let\onelinecomment=\empty
  \let\commentbegin=\empty
  \let\commentend=\empty}

\visiblecomments

\newlength{\blanklineskip}
\setlength{\blanklineskip}{0.66084ex}

\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
\let\hspre\empty
\let\hspost\empty
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

\EndFmtInput
\makeatother
%
%
%
%
%
%
% This package provides two environments suitable to take the place
% of hscode, called "plainhscode" and "arrayhscode". 
%
% The plain environment surrounds each code block by vertical space,
% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
% similar to formulas. Note that if these dimensions are changed,
% the spacing around displayed math formulas changes as well.
% All code is indented using \leftskip.
%
% Changed 19.08.2004 to reflect changes in colorcode. Should work with
% CodeGroup.sty.
%
\ReadOnlyOnce{polycode.fmt}%
\makeatletter

\newcommand{\hsnewpar}[1]%
  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}

% can be used, for instance, to redefine the code size, by setting the
% command to \small or something alike
\newcommand{\hscodestyle}{}

% The command \sethscode can be used to switch the code formatting
% behaviour by mapping the hscode environment in the subst directive
% to a new LaTeX environment.

\newcommand{\sethscode}[1]%
  {\expandafter\let\expandafter\hscode\csname #1\endcsname
   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}

% "compatibility" mode restores the non-polycode.fmt layout.

\newenvironment{compathscode}%
  {\par\noindent
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed\)%
   \par\noindent
   \ignorespacesafterend}

\newcommand{\compaths}{\sethscode{compathscode}}

% "plain" mode is the proposed default.
% It should now work with \centering.
% This required some changes. The old version
% is still available for reference as oldplainhscode.

\newenvironment{plainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newenvironment{oldplainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

% Here, we make plainhscode the default environment.

\newcommand{\plainhs}{\sethscode{plainhscode}}
\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
\plainhs

% The arrayhscode is like plain, but makes use of polytable's
% parray environment which disallows page breaks in code blocks.

\newenvironment{arrayhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\parray}%
  {\endparray\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newcommand{\arrayhs}{\sethscode{arrayhscode}}

% The mathhscode environment also makes use of polytable's parray 
% environment. It is supposed to be used only inside math mode 
% (I used it to typeset the type rules in my thesis).

\newenvironment{mathhscode}%
  {\parray}{\endparray}

\newcommand{\mathhs}{\sethscode{mathhscode}}

% texths is similar to mathhs, but works in text mode.

\newenvironment{texthscode}%
  {\(\parray}{\endparray\)}

\newcommand{\texths}{\sethscode{texthscode}}

% The framed environment places code in a framed box.

\def\codeframewidth{\arrayrulewidth}
\RequirePackage{calc}

\newenvironment{framedhscode}%
  {\parskip=\abovedisplayskip\par\noindent
   \hscodestyle
   \arrayrulewidth=\codeframewidth
   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
   \hline\framedhslinecorrect\\{-1.5ex}%
   \let\endoflinesave=\\
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \framedhslinecorrect\endoflinesave{.5ex}\hline
   \endtabular
   \parskip=\belowdisplayskip\par\noindent
   \ignorespacesafterend}

\newcommand{\framedhslinecorrect}[2]%
  {#1[#2]}

\newcommand{\framedhs}{\sethscode{framedhscode}}

% The inlinehscode environment is an experimental environment
% that can be used to typeset displayed code inline.

\newenvironment{inlinehscode}%
  {\(\def\column##1##2{}%
   \let\>\undefined\let\<\undefined\let\\\undefined
   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
   \def\fromto##1##2##3{##3}%
   \def\nextline{}}{\) }%

\newcommand{\inlinehs}{\sethscode{inlinehscode}}

% The joincode environment is a separate environment that
% can be used to surround and thereby connect multiple code
% blocks.

\newenvironment{joincode}%
  {\let\orighscode=\hscode
   \let\origendhscode=\endhscode
   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
   %\let\SaveRestoreHook=\empty
   %\let\ColumnHook=\empty
   %\let\resethooks=\empty
   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
  {\origendhscode
   \global\let\hscode=\orighscode
   \global\let\endhscode=\origendhscode}%

\makeatother
\EndFmtInput
%
%
\ReadOnlyOnce{agda.fmt}%

\RequirePackage[T1]{fontenc}
\RequirePackage[utf8x]{inputenc}
\RequirePackage{ucs}
\RequirePackage{amsfonts}

\providecommand\mathbbm{\mathbb}

% TODO: Define more of these ...
\DeclareUnicodeCharacter{737}{\textsuperscript{l}}
\DeclareUnicodeCharacter{8718}{\ensuremath{\blacksquare}}
\DeclareUnicodeCharacter{8759}{::}
\DeclareUnicodeCharacter{9669}{\ensuremath{\triangleleft}}
\DeclareUnicodeCharacter{8799}{\ensuremath{\stackrel{\scriptscriptstyle ?}{=}}}
\DeclareUnicodeCharacter{10214}{\ensuremath{\llbracket}}
\DeclareUnicodeCharacter{10215}{\ensuremath{\rrbracket}}

% TODO: This is in general not a good idea.
\providecommand\textepsilon{$\epsilon$}
\providecommand\textmu{$\mu$}

%Actually, varsyms should not occur in Agda output.

% TODO: Make this configurable. IMHO, italics doesn't work well
% for Agda code.

\renewcommand\Varid[1]{\mathord{\textsf{#1}}}
\let\Conid\Varid
\newcommand\Keyword[1]{\textsf{\textbf{#1}}}
\EndFmtInput

\begin{document}

\title{Indexed Containers}
\author{Thorsten Altenkirch 
   \and Neil Ghani 
   \and Peter Hancock 
   \and Conor McBride 
   \and Peter Morris}
\date{\today}

\maketitle

\begin{abstract}

Blah

\end{abstract}

\section{Introduction}

\section{Background}

\subsection{Type Theory}

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\anonymous \dot{\rightarrow}\anonymous \;\mathbin{:}\;\{\mskip1.5mu \Conid{I}\;\mathbin{:}\;\Conid{Set}\mskip1.5mu\}\;\to \;(\Conid{A}\;\Conid{B}\;\mathbin{:}\;\Conid{I}\;\to \;\Conid{Set})\;\to \;\Conid{Set}{}\<[E]%
\\
\>[B]{}\anonymous \dot{\rightarrow}\anonymous \;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;\Conid{A}\;\Conid{B}\;\mathrel{=}\;(\Varid{i}\;\mathbin{:}\;\Conid{I})\;\to \;\Conid{A}\;\Varid{i}\;\to \;\Conid{B}\;\Varid{i}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\subsection{Containers}

\section{Indexed Functors}

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{6}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{record}\;\Conid{IFunc}\;(\Conid{I}\;\mathbin{:}\;\Conid{Set})\;\mathbin{:}\;\Conid{Set}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{field}{}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{}\Varid{obj}\;\mathbin{:}\;(\Conid{A}\;\mathbin{:}\;\Conid{I}\;\to \;\Conid{Set})\;\to \;\Conid{Set}{}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{}\Varid{mor}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{A}\;\Conid{B}\mskip1.5mu\}\;\to \;(\Conid{A}\;\dot{\rightarrow}\;\Conid{B})\;\to \;\Varid{obj}\;\Conid{A}\;\to \;\Varid{obj}\;\Conid{B}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{12}{@{}>{\hspre}c<{\hspost}@{}}%
\column{12E}{@{}l@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{28}{@{}>{\hspre}l<{\hspost}@{}}%
\column{48}{@{}>{\hspre}l<{\hspost}@{}}%
\column{61}{@{}>{\hspre}l<{\hspost}@{}}%
\column{68}{@{}>{\hspre}l<{\hspost}@{}}%
\column{71}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\eta^{F}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;\rightarrow\;\Conid{I}\;\rightarrow\;\Conid{IFunc}\;\Conid{I}{}\<[E]%
\\
\>[B]{}\eta^{F}\;\Varid{i}\;\mathrel{=}\;\Keyword{record}\;\{\mskip1.5mu \Varid{obj}\;\mathrel{=}\;\lambda\;\Conid{A}\;\rightarrow\;\Conid{A}\;\Varid{i};\Varid{mor}\;\mathrel{=}\;\lambda\;\Varid{f}\;\rightarrow\;\Varid{f}\;\Varid{i}\mskip1.5mu\}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\_\bind ^{F}\_\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\mskip1.5mu\}\;\rightarrow\;(\Conid{I}\;\rightarrow\;\Conid{IFunc}\;\Conid{J})\;\rightarrow\;\Conid{IFunc}\;\Conid{I}\;\rightarrow\;\Conid{IFunc}\;\Conid{J}{}\<[E]%
\\
\>[B]{}\Conid{H}\;\bind ^{F}\;\Conid{F}\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{record}\;{}\<[12]%
\>[12]{}\{\mskip1.5mu {}\<[12E]%
\>[15]{}\Varid{obj}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Conid{A}\;{}\<[28]%
\>[28]{}\rightarrow\;\!\;\Conid{F}\;(\lambda\;\Varid{i}\;{}\<[48]%
\>[48]{}\rightarrow\;\!\;{}\<[61]%
\>[61]{}(\Conid{H}\;\Varid{i})\;{}\<[68]%
\>[68]{}\Conid{A}{}\<[71]%
\>[71]{}){}\<[E]%
\\
\>[12]{};{}\<[12E]%
\>[15]{}\Varid{mor}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Varid{f}\;{}\<[28]%
\>[28]{}\rightarrow\;\!\;\Conid{F}\;(\lambda\;\Varid{i}\;{}\<[48]%
\>[48]{}\rightarrow\;\!\;{}\<[61]%
\>[61]{}(\Conid{H}\;\Varid{i})\;{}\<[68]%
\>[68]{}\Varid{f}{}\<[71]%
\>[71]{})\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{IFunc}^{\star}\;\mathbin{:}\;(\Conid{I}\;\Conid{J}\;\mathbin{:}\;\Conid{Set})\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\mathrel{=}\;\Conid{J}\;\rightarrow\;\Conid{IFunc}\;\Conid{I}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{obj}^{\star}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\mskip1.5mu\}\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;(\Conid{I}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{J}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Varid{obj}^{\star}\;\Conid{F}\;\Conid{A}\;\Varid{j}\;{}\<[16]%
\>[16]{}\mathrel{=}\;\!\;(\Conid{F}\;\Varid{j})\;\Conid{A}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{mor}^{\star}\;\mathbin{:}\;{}\<[10]%
\>[10]{}\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\;\Conid{A}\;\Conid{B}\mskip1.5mu\}\;(\Conid{F}\;\mathbin{:}\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J})\;\rightarrow\;\Conid{A}\;\dot{\rightarrow}\;\Conid{B}\;\rightarrow\;\Varid{obj}^{\star}\;\Conid{F}\;\Conid{A}\;\dot{\rightarrow}\;\Varid{obj}^{\star}\;\Conid{F}\;\Conid{B}{}\<[E]%
\\
\>[B]{}\Varid{mor}^{\star}\;\Conid{F}\;\Varid{m}\;\Varid{j}\;{}\<[14]%
\>[14]{}\mathrel{=}\;\!\;(\Conid{F}\;\Varid{j})\;\Varid{m}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{12}{@{}>{\hspre}c<{\hspost}@{}}%
\column{12E}{@{}l@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{40}{@{}>{\hspre}l<{\hspost}@{}}%
\column{91}{@{}>{\hspre}c<{\hspost}@{}}%
\column{91E}{@{}l@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Sigma^{F}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\;\Conid{K}\mskip1.5mu\}\;\rightarrow\;(\Conid{J}\;\rightarrow\;\Conid{K})\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{K}{}\<[E]%
\\
\>[B]{}\Sigma^{F}\;\{\mskip1.5mu \Conid{J}\;\mathrel{=}\;\Conid{J}\mskip1.5mu\}\;\Varid{f}\;\Conid{F}\;\Varid{k}\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{record}\;{}\<[12]%
\>[12]{}\{\mskip1.5mu {}\<[12E]%
\>[15]{}\Varid{obj}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Conid{A}\;\rightarrow\;\Sigma\;\Conid{J}\;\lambda\;\Varid{j}\;\rightarrow\;\Varid{f}\;\Varid{j}\;\equiv\;\Varid{k}\;\times\;\!\;\Conid{F}\;\Conid{A}\;\Varid{j}{}\<[E]%
\\
\>[12]{};{}\<[12E]%
\>[15]{}\Varid{mor}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Varid{m}\;\rightarrow\;\Varid{<}\;\pi_{0},{}\<[40]%
\>[40]{}\Varid{<}\;\pi_{0}\;\ensuremath{\mbox{$\circ$}}\;\pi_{1},\!\;\Conid{F}\;\Varid{m}\;\anonymous \;\ensuremath{\mbox{$\circ$}}\;\pi_{1}\;\ensuremath{\mbox{$\circ$}}\;\pi_{1}\;\Varid{>}\;\Varid{>}{}\<[91]%
\>[91]{}\mskip1.5mu\}{}\<[91E]%
\\[\blanklineskip]%
\>[B]{}\Pi^{F}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\;\Conid{K}\mskip1.5mu\}\;\rightarrow\;(\Conid{J}\;\rightarrow\;\Conid{K})\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{K}{}\<[E]%
\\
\>[B]{}\Pi^{F}\;\{\mskip1.5mu \Conid{J}\;\mathrel{=}\;\Conid{J}\mskip1.5mu\}\;\Varid{f}\;\Conid{F}\;\Varid{k}\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{record}\;{}\<[12]%
\>[12]{}\{\mskip1.5mu {}\<[12E]%
\>[15]{}\Varid{obj}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Conid{A}\;\rightarrow\;(\Varid{j}\;\mathbin{:}\;\Conid{J})\;\rightarrow\;\Varid{f}\;\Varid{j}\;\equiv\;\Varid{k}\;\rightarrow\;\!\;\Conid{F}\;\Conid{A}\;\Varid{j}{}\<[E]%
\\
\>[12]{};{}\<[12E]%
\>[15]{}\Varid{mor}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Varid{m}\;\Varid{f}\;\Varid{j}\;\Varid{p}\;\rightarrow\;\!\;\Conid{F}\;\Varid{m}\;\Varid{j}\;(\Varid{f}\;\Varid{j}\;\Varid{p})\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}c<{\hspost}@{}}%
\column{11E}{@{}l@{}}%
\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
\column{46}{@{}>{\hspre}l<{\hspost}@{}}%
\column{50}{@{}>{\hspre}l<{\hspost}@{}}%
\column{54}{@{}>{\hspre}l<{\hspost}@{}}%
\column{59}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\_[\mskip1.5mu \_\mskip1.5mu]^{F}\;\mathbin{:}\;\forall{}\;\{\mskip1.5mu \Conid{I}\;\Conid{J}\mskip1.5mu\}\;\rightarrow\;\Conid{IFunc}\;(\Conid{I}\;\uplus\;\Conid{J})\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;{}\<[50]%
\>[50]{}\Conid{IFunc}\;\Conid{I}{}\<[E]%
\\
\>[B]{}\Conid{F}\;[\mskip1.5mu \;\Conid{G}\;\mskip1.5mu]^{F}\;{}\<[12]%
\>[12]{}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{record}\;{}\<[11]%
\>[11]{}\{\mskip1.5mu {}\<[11E]%
\>[14]{}\Varid{obj}\;{}\<[19]%
\>[19]{}\mathrel{=}\;{}\<[22]%
\>[22]{}\lambda\;\Conid{A}\;{}\<[27]%
\>[27]{}\rightarrow\;\!\;\Conid{F}\;[\mskip1.5mu \Conid{A}{}\<[46]%
\>[46]{},\!\;{}\<[54]%
\>[54]{}\Conid{G}\;\Conid{A}{}\<[59]%
\>[59]{}\mskip1.5mu]{}\<[E]%
\\
\>[11]{};{}\<[11E]%
\>[14]{}\Varid{mor}\;{}\<[19]%
\>[19]{}\mathrel{=}\;{}\<[22]%
\>[22]{}\lambda\;\Varid{f}\;{}\<[27]%
\>[27]{}\rightarrow\;\!\;\Conid{F}\;[\mskip1.5mu \Varid{f}{}\<[46]%
\>[46]{},\!\;{}\<[54]%
\>[54]{}\Conid{G}\;\Varid{f}{}\<[59]%
\>[59]{}\mskip1.5mu]\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\section{Indexed containers}

\section{Initial Algebras of Indexed Containers}

\section{Strictly Positive Types}

\section{Conclusions}

\end{document}